### 2021-11-16 newvariant.tex
almost there?

We explored how the generation-interval distribution shapes the link between relative strength and speed of an invading pathogen variant.
Longer generation intervals generally lead to higher relative strength for a given relative speed (and conversely, lower relative speed for a given relative strength); these relationships are reversed when incidence is decreasing.
Neglecting potential differences in the mean generation intervals between the variant and the wild type can bias estimates of the relative strength from incidence data;
these biases may be assessed by considering whether estimates of relative strength appear to vary systematically with the direction of changes in the incidence of infections caused by the variant.
Finally, differences in generation intervals can also lead to different conclusions about the effectiveness of interventions.
If the variant has longer generation intervals than the wild type, speed-like interventions will be relatively more effective than naive estimates would suggest. 
Conversely, strength-like intervention will be relatively more effective if the variant has shorter generation intervals.

>>>>

We explored how the generation-interval distribution shapes the link between relative strength and speed.
While the role of generation-interval distributions in linking epidemic strength and speed has been well-established, using the generation-interval-based framework allowed us to obtain imporant insight: the relative strength and speed of new variants also depend on epidemic dynamics of the prior dominant lineages as well as the type of control measures (i.e., strength- or speed-like) that limit their spread.
For example, under a constant-strength intervention that reduces the transmission rate of the wild type and a variant by the same amount (and therefore, keeping the relative strength constant), the relative speed of the variant is sensitive to

Longer generation intervals generally lead to higher relative strength for a given relative speed (and conversely, lower relative speed for a given relative strength), but these relationships are reversed when incidence is decreasing.

Neglecting potential differences in the mean generation intervals between the variant and the wild type can bias estimates of the relative strength from incidence data;
these biases may be assessed by considering whether estimates of relative strength appear to vary systematically with the direction of changes in the incidence of infections caused by the variant.
Finally, differences in generation intervals can also lead to different conclusions about the effectiveness of interventions.
If the variant has longer generation intervals than the wild type, speed-like interventions will be relatively more effective than naive estimates would suggest. 
Conversely, strength-like intervention will be relatively more effective if the variant has shorter generation intervals.
